---
title: "Final Project Prelim Analysis"
author: "Kate Jones"
date: "10/11/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,cache=TRUE)
```

# First attempt at plot level data: Coweeta LTER

***My first attempt at finding plot level data for image classification lead me to the Coweeta LTER in western NC. The data were accessible and had a kml file ready to go, but unfortunatley, the plot dimensions, nor the corner coordinates were made clear. Below is the preliminary exploration of plot data conducted before realizing the georeferencing was not up to snuff. ***

```{r}
library(data.table)
# GOAL: figure out which plots within Coweeta data have pure species information
setwd("C:\\Users\\The Brain\\OneDrive - North Carolina State University\\F2019 - Classes\\GIS 713\\Final-Project")

#read in csv file of coweeta data from all plots
cow_trees<-as.data.frame(read.csv(".\\Tree_Data_1_1155.csv",header=TRUE,row.names=NULL,stringsAsFactors = FALSE))
#clean up the dataframe
cow_trees<-cow_trees[-c(1:4),]
#renme the columns
colnames(cow_trees) = c("date","site","plot","species","dbh","status","comments")
#turn dbh to numeric 
cow_trees$dbh<-as.numeric(cow_trees$dbh)

#turn this to a data.table
cow_dt<-as.data.table(cow_trees)

#after looking at the kml files, I choose a test site that would have plots entirely in the forest, so the imagery would be entirely forest and not other landcover types
fall_site<-as.data.table(cow_dt[cow_dt$site=="FALL",])

#now i'm interested in finding plots that are almost entirely one species
byplot_spec<-fall_site[dbh>10,species,by=plot]

#specifically, i'm likely interested in larger trees because of their presence in the canopy
byplot_numspec<-fall_site[dbh>10,.N,by=list(plot,site)]

#filtering out for live trees
siteandplot<-cow_dt[dbh>10|status=="alive",no_of_spec := length(unique(species)),by=list(plot,site)]

#this returns the number of species (of living trees) tallied for each plot at this site
nonas<-na.omit(siteandplot,cols="no_of_spec")
setorder(nonas, -no_of_spec)

```

# Vegbank Data

***I quickly tabeled the Coweeta data after the suggestion to use the VegBank data (http://vegbank.org/vegbank/index.jsp).***

***The accuracy differs across Vegbank plots, but my criteria for searching were:*** 

1. Greater than 85% canopy cover for the SOI
  + white oak
  + red maple
2. Classified as accurate to 10m
3. Exact latitude and longitude (centroid) given for a rectangular plot


***For my oreliminary analysis, I did not download any VegBank data. I simply used their advanced query system to find plots that meet the criteria above. I then manually created polygons in R with the appropriate centroid and extent.***

***One serious road block for this analysis could be that the data collection years for VegBank data varies widely. The years for data collection might not correspond well to the much newer Sentinel-2 data. Although, I am trying classify late successional forest, so given there was no major disturbance, the species information from VegBank could still be accurate.***

#Creating plot polygons from VegBank data: White Oak

```{r, echo=FALSE}

library(rgdal)
library(sp)

#plot no. = 042-05-0629
#store centroid coordinate for plot in western NC with canopy 85% white oak 
ll_x<- -82.580218036 
ll_y<- 35.205450602

#store in object to pass to SPointsDF coords argument
ll_coords<-cbind(ll_x,ll_y)
#make a df to pass to SpointsDF df argument
df<-as.data.frame(ll_coords)
#input projection
ll <-"+init=epsg:4326"
#what we want the projection to be
utm<-"+proj=utm +zone=17 +ellps=WGS84 +datum=WGS84 +units=m +no_defs"

#create centroid SPDF
wo_centroid<-SpatialPointsDataFrame(ll_coords,df)
#set projection for centroid in input form (latlong)
proj4string(wo_centroid)<-CRS(ll)
#transform centroid to utm
wo_centroid_utm <- spTransform(wo_centroid, CRS(utm))

#then get reprojected centroid
utm_centroid<-as.data.frame(wo_centroid_utm@coords)
colnames(utm_centroid)<-c("x","y")

#get centroids in utm
utm_x<-utm_centroid$x 
utm_y<-utm_centroid$y

# radius in meters, the plots are 1000 m2
radius <- sqrt(1000)/2

#make twice the size for extracting pixels
big_radius<-sqrt(1000)

# define the plot edges based upon the plot radius
yPlus <- utm_y+radius
xPlus <- utm_x+radius
yMinus <- utm_y-radius
xMinus <- utm_x-radius

#make bigger box
yplus_big<-utm_y+big_radius
xplus_big<-utm_x+big_radius
yminus_big<-utm_y-big_radius
xminus_big<-utm_x-big_radius

utm_bbox = matrix(c(xMinus,yPlus,# NW corner
                xPlus, yPlus,# NE corner
                xPlus,yMinus,# SE corner
                xMinus,yMinus,# SW corner
                xMinus,yPlus), # NW corner again - close ploygon
             ncol = 2, byrow = TRUE)

big_utm_bbox= matrix(c(xminus_big,yplus_big,# NW corner
                xplus_big, yplus_big,# NE corner
                xplus_big,yminus_big,# SE corner
                xminus_big,yminus_big,# SW corner
                xminus_big,yplus_big), # NW corner again - close ploygon
                ncol = 2, byrow = TRUE)


P1 = Polygon(utm_bbox)
wo_utm_bbox = SpatialPolygons(list(Polygons(list(P1), ID = "a")), proj4string=CRS(utm))
#plot(wo_utm_bbox, axes = TRUE)
#points(utm_centroid$x,utm_centroid$y)

#create a bigger box for extracting pixels
P2 = Polygon(big_utm_bbox)
big_wo_utm_bbox = SpatialPolygons(list(Polygons(list(P2), ID = "b")), proj4string=CRS(utm))

```

# Reading in Sentinel Data

```{r}

library(utils)
library(raster)

zipfile<-"C:\\Users\\The Brain\\Downloads\\S2B_MSIL2A_20190815T160829_N0213_R140_T17SLV_20190815T203237.zip"

unzip(zipfile, exdir = ".")

path_to_rast<-"C:\\Users\\The Brain\\OneDrive - North Carolina State University\\F2019 - Classes\\GIS 713\\Final-Project\\sent\\GRANULE\\L2A_T17SLV_A012750_20190815T161632\\IMG_DATA\\R10m"

raster_data<-list.files(path=path_to_rast,full.names = TRUE)

s <- raster::stack(raster_data)
stack<-s[[2:5]]

stack.sub <- crop(stack, extent(wo_utm_bbox))
#clip_rasts <- raster::mask(wo_utm_bbox, stack)

b02 <- raster(raster_data[2])#produces raster layer
b03 <- raster(raster_data[3])#produces sp object
b04 <- readGDAL(raster_data[4])#originally used rgdal because i was worried about about sentinal file types
b08 <- readGDAL(raster_data[5])#produces sp object

b03_crop <- raster::intersect(b03, big_wo_utm_bbox)

plot(b03_crop)
plot(wo_utm_bbox, add=TRUE)
points(utm_centroid)
plot(big_wo_utm_bbox,add=TRUE)


plot(mask(b03_crop, big_wo_utm_bbox))
```

# Creating plot polygon from VegBank data: Red Maple

```{r}

#plot no. = 017-01-0026
#store centroid coordinate for plot in western NC with canopy 85% red maple 
rm_ll_x<- -77.036533569
rm_ll_y<- 35.861254385

#store in object to pass to SPointsDF coords argument
ll_rm_coords<-cbind(rm_ll_x,rm_ll_y)
#make a df to pass to SpointsDF df argument
rm_df<-as.data.frame(ll_rm_coords)

#create centroid SPDF
rm_centroid<-SpatialPointsDataFrame(ll_rm_coords,rm_df)
#set projection for centroid in input form (latlong)
proj4string(rm_centroid)<-CRS(ll)
#transform centroid to utm
rm_centroid_utm <- spTransform(rm_centroid, CRS(utm))

#then get reprojected centroid
rm_utm_centroid<-as.data.frame(rm_centroid_utm@coords)
colnames(rm_utm_centroid)<-c("x","y")

#get centroids in utm
rm_utm_x<-rm_utm_centroid$x 
rm_utm_y<-rm_utm_centroid$y


# define the plot edges based upon the plot radius
yPlus <- rm_utm_y+radius
xPlus <- rm_utm_x+radius
yMinus <- rm_utm_y-radius
xMinus <- rm_utm_x-radius

#make bigger box
yplus_big<-rm_utm_y+big_radius
xplus_big<-rm_utm_x+big_radius
yminus_big<-rm_utm_y-big_radius
xminus_big<-rm_utm_x-big_radius

r_utm_bbox = matrix(c(xMinus,yPlus,# NW corner
                xPlus, yPlus,# NE corner
                xPlus,yMinus,# SE corner
                xMinus,yMinus,# SW corner
                xMinus,yPlus), # NW corner again - close ploygon
             ncol = 2, byrow = TRUE)

big_rm_utm_bbox= matrix(c(xminus_big,yplus_big,# NW corner
                xplus_big, yplus_big,# NE corner
                xplus_big,yminus_big,# SE corner
                xminus_big,yminus_big,# SW corner
                xminus_big,yplus_big), # NW corner again - close ploygon
                ncol = 2, byrow = TRUE)


P1 = Polygon(r_utm_bbox)
rm_utm_bbox = SpatialPolygons(list(Polygons(list(P1), ID = "a")), proj4string=CRS(utm))

#create a bigger box for extracting pixels
P2 = Polygon(big_rm_utm_bbox)
big_rm_utm_bbox = SpatialPolygons(list(Polygons(list(P2), ID = "b")), proj4string=CRS(utm))

b03_crop1<- raster::intersect(b03, big_rm_utm_bbox)

plot(b03_crop1)
plot(rm_utm_bbox, add=TRUE)
points(rm_utm_centroid)
plot(big_rm_utm_bbox,add=TRUE)


plot(mask(b03_crop, big_wo_utm_bbox))

```

